% !TeX spellcheck = en_GB
% encoding: utf8
% !TEX encoding = utf8
% !TEX program = pdflatex
% !BIB program = biber

\documentclass[11pt,oneside,a4paper]{article}

\usepackage[cmex10]{amsmath}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage[linesnumbered,ruled,vlined,onelanguage]{algorithm2e}
\usepackage{algorithmic}
\usepackage{mathtools}

\usepackage{color, colortbl}
\usepackage{footmisc}

\usepackage{ifthen}

%%%%%
%%  Comments 
%%%%%
\usepackage{xargs}
\newcommand{\Fig}[1]{Fig.~\ref{#1}}

\usepackage{fourier} 
\usepackage{array}
\usepackage{makecell}
\usepackage{csvsimple}
\usepackage{booktabs}

\renewcommand\theadalign{bc}
\renewcommand\theadfont{\bfseries}
\renewcommand\theadgape{\Gape[4pt]}
\renewcommand\cellgape{\Gape[4pt]}
\renewcommand{\labelenumii}{\theenumii}
\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}
\usepackage{tabularx}

\newcommand{\twc}[1]{\textcolor{red}{\textbf{TW}: #1}}
\usepackage{hyperref} %<--- Load after everything else
\hypersetup{
	colorlinks = true,
}

%%%%%
%%   Definicja wymagań
%%%%
\usepackage[backend=biber,maxnames=10,style=numeric,sorting=nty,abbreviate=false,giveninits=true,language=english]{biblatex}
\addbibresource{meros.bib}

\newcommand{\req}[1]{\textbf{\texttt{R#1}}}
\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{//#1\\}}
\begin{document}

\title{MeROS: SysML-based Metamodel for ROS-based Systems - version 1.1 - reference manual}
\author{MeROS developers group - https://github.com/twiniars/MeROS \\ tomasz.winiarski@pw.edu.pl}
\date{\today}
\maketitle
	\begin{abstract}
		The complexity of today's robot control systems implies difficulty in developing them efficiently and reliably. Systems engineering (SE) and frameworks come to help. The framework metamodels are needed to support the standardisation and correctness of the created application models. MeROS is a~metamodel for ROS, which addresses the running system and developer workspace. The ROS comes in two versions: ROS~1 and ROS~2. The metamodel includes both versions. In particular, the latest ROS~1 concepts are considered, such as nodelet, action, and metapackage. An essential addition to the original ROS concepts is the grouping of these concepts, which provides an opportunity to illustrate the system's decomposition and varying degrees of detail in its presentation. The metamodel is derived from the requirements and verified on the practical examples. The matter is described in a standardised way in SysML (Systems Modeling Language). Hence, common development tools that support SysML can help develop robot controllers in the spirit of SE.
	\end{abstract}
	\maketitle
	\section*{Important citation notice}
\textbf{If you are to use MeROS in your papers, please first cite the IEEE ACCESS  article~\cite{meros-access}, where the initial version of the MeROS is presented. You can also refer to MeROS project page \cite{meros-www} and mention the actual MeROS version you are using.}

\input{tex/introduction}
\input{tex/metamodel-requirements}
\input{tex/meros-metamodel}
\input{tex/meros-application}

\AtNextBibliography{\small}
\printbibliography
	
\end{document}
